### 2015-09-03 chapters/conclusion.tex
basic edits to intro and conclusion

In Chapter \ref{chap:v4046_structure} we 
used the SMA to study the protoplanetary disk 
around the close binary V4046 Sgr.  This disk featured a narrow ring of large mm-sized grains along with
an extended gas rich disk.  Detailed modeling of these features demonstrated that they were 
consistent with the viscous spreading of the gas and the concentration of the dust via pressure 
drag forces.  This disk is one published example among a few and more similar systems continue to be
identified \citep{isella07,panic09,andrews12,degregorio-monsalvo13,ke14}.  A related phenomena are the 
severely asymmetric dust  ``horse-shoes'' that have been  observed at multiple wavelengths
\citep{vandermarel13,fukagawa13,isella13,casassus15,marino15}.  Thought to be perhaps generated by gigantic 
3D Rossby wave instabilities \citep{regaly13}, these are another striking example of the interplay between
the gas and dust in protoplanetary disks.  A particular exciting example are the narrow gaps in the dust continuum
of HL Tau (see Figure \ref{fig:hltau}. This striking image was obtained using ALMA baselines as long as 15\,km 
and possible causes include
planet-disk interactions and condensation fronts \citep{brogan14,zhang15}.

\begin{figure}
\plotone{chapters/HLTau_nrao.jpg}
\caption{This ALMA observation of the HL Tau disk revealed multiple rings and gaps \citep{brogan15}.}
\label{fig:hltau}
\end{figure}

@@

To conclude our work on protoplanary disks, we discussed in Chapter \ref{chap:radial_flows} observational 
indicators of 
radial gas flows within the holes of transition disks.  \citet{casassus15} have studied the gas kinematics for
the curious case of HD142527 using the CO $J = 6-5$ line and found that their data is consistent with
accretion reaching close to free-fall velocities.  The gas surface density of transition disks have also 
been recently studied using observations taking at sub-millimeter wavelengths 
\citep{bruderer14,zhang14,canovas15,perez15,vandermarel15} and will help determine how these structures are formed
\citep{bruderer13}.  Molecular disk winds, outflows, and binary systems are all recently observed 
with interesting gas kinematics \citep{klaasen13,dutrey14,williams14,salyk14,czekala15}.  Looking to the 
future, well resolved, highly sensitive observations of spectral lines may constrain the strength of MRI-drive
turbulence \citep{simon15}, query the active sites of planet formation \citep{kleeves15,ober15}, or reveal
circumplanetary disks \citep{perez15}.

The ground-breaking Event Horizon Telescope continues to report exciting new science with every
campaign from the early detections of AGN at mm wavelengths \citep{doeleman05} to structures
at the event horizon of a supermassive black hole \citep{doeleman08,doeleman09,doeleman13}.
The 2015 campaign represented a new scale of operation, incorporating more telescopes and
technologies.  In Chapter \ref{chap:aphids} we reported on the status of the APHIDS software which
will post-process the phased SMA data and make the sampling rate compatible with the rest of the EHT
network.  This problem is not limited to the SMA and we hope that the development done with APHIDS will
prove useful for other sites and VLBI experiments.

This thesis explores
ence that resolution at these wavelengths can provide and only an exploration for the 
hard and promising work ahead.

>>>>

In Chapter \ref{chap:v4046sgr_structure} we 
used the SMA to study the protoplanetary disk 
around the close binary V4046 Sgr.  This disk featured a narrow ring of large mm-sized grains along with
an extended gas rich disk.  Detailed modeling of these features demonstrated that they were 
consistent with the viscous spreading of the gas and the concentration of the dust via pressure 
drag forces.  This disk is one published example among a few and more similar systems continue to be
identified \citep{isella07,panic09,andrews12,degregorio-monsalvo13,zhang14}.  A related phenomenon are the 
severely asymmetric dust  ``horse-shoe'' disks that have been observed at multiple wavelengths
\citep{vandermarel13,fukagawa13,isella13,casassus15,marino15}.  Thought to be perhaps generated by gigantic 
3D Rossby wave instabilities \citep{regaly12}, these are another striking example of the interplay between
the gas and dust in protoplanetary disks.  Another particular exciting example is HL Tau where recent
observations have revealed narrow gaps in the dust 
continuum (see Figure \ref{fig:hltau}. This image was obtained using ALMA baselines as long as 15\,km 
and possible causes include
planet-disk interactions and condensation fronts \citep{brogan15,zhang15}.

\begin{figure}
\centering
\includegraphics[width=0.5\linewidth]{chapters/HLTau_nrao.jpg}
\caption{This ALMA observation of the HL Tau disk revealed multiple rings and gaps \citep{brogan15}.}
\label{fig:hltau}
\end{figure}

@@

To conclude our work on protoplanary disks, we discussed in Chapter \ref{chap:radial_flows} observational 
indicators of 
radial gas flows within the holes of transition disks.  \citet{casassus15} have studied the gas kinematics for
the curious case of HD142527 using the CO $J$ = $6$-$5$ line and found that their data is consistent with
accretion reaching close to free-fall velocities.  The gas surface density of transition disks have also 
been recently studied using observations taking at sub-millimeter wavelengths 
\citep{bruderer14,zhang14,canovas15,perez15,vandermarel15} and will help determine how these structures are formed
\citep{bruderer13}.  Molecular disk winds, outflows, and binary systems are all recently observed 
with interesting gas kinematics \citep{klaassen13,dutrey14,williams14,salyk14,czekala15}.  Looking to the 
future, well resolved, highly sensitive observations of spectral lines may constrain the strength of MRI-drive
turbulence \citep{simon15}, query the active sites of planet formation \citep{cleeves15,ober15}, or reveal
circumplanetary disks \citep{perez15}.

The ground-breaking Event Horizon Telescope continues to report exciting, new science with every
campaign from the early detections of AGN at mm wavelengths \citep{doeleman05} to structures
at the event horizon of a supermassive black hole \citep{doeleman08,doeleman12}.
The 2015 campaign represented a new scale of operations, incorporating more telescopes and
fielding new technologies.  In Chapter \ref{chap:aphids} we reported on the status of the APHIDS software which
will post-process the phased SMA data and make the sampling rate compatible with the rest of the EHT
network.  This problem is not limited to the SMA and we hope that the development done with APHIDS will
prove useful for other sites and VLBI experiments.

### 2015-09-03 chapters/introduction.tex
basic edits to intro and conclusion

Millimeter interferometry is intimately tied to the earlier development of techniques and technologies 
used at longer wavelengths.  The essential operation combines the signals from pairs of telescopes to produce a 
measure that is sensitive to some angular scale on the sky.  This scale shrinks as the distance
between the two elements grows.  By assembling an array with multiple 
telescopes, an observer can accumulate information about many spatial scales and reconstruct an image.  However,
the information contained in this image will be limited to the scales originally measured by the interferometer.
Images with the best fidelity come from arrays that have elements at a wide variety of 
distances to each other and the highest resolution comes from the most widely separated telescopes.

@@

Each pair of antennas in an array forms a baseline vector (measured in wavelengths) with length $D_\lambda$ and 
components $(u,v,w)$.  The signal voltages, $V_1(t)$ and $V_2(t)$, measured by the individual 
elements are correlated against each other:
\begin{equation}
r(\tau) = \lim_{\tau \to \infty} \frac{1}{2T} \int_{-T}^T V_1(t) V_2^\ast(t-\tau) d\tau.
\end{equation}
This operation is often done on site with data streamed directly to a correlator and the resulting 
measurement, or visibility, is stored on disk.  However, for Very Long Baseline Interferometry (VLBI), the 
radio telescopes can be separated by entire continents and real-time correlation is not usually feasible.  For 
these experiments, the signal voltages must be recorded and the visibilities calculated at some later time.  The 
Event Horizon Telescope (EHT) is a leader in the developing field of millimeter VLBI and relies on existing 
observatories.  The difficulty of observing at these short wavelengths increases as the atmosphere becomes 
noiser, time variable, and more opaque.  Consequently, facilities that can operate at these wavelengths are a 
precious resource.

The Submillimeter Array \citep[SMA][]{ho04} is one such world class array.  Commisioned in 2003, the SMA is 
located on the top of Maunakea on the Big Island of Hawaii at an altitude of 13,386 feet.
It is composed of eight 6 meter diameter dishes which, like many interferometers, can be moved into different 
configurations so that there are elements as close as 15 meters or as far as 500 meters (see Figure 
\ref{fig:sma_alma}).  Its recievers can operate at wavelengths from 430\,microns to 2\,millimeters and 
historically have provided up to 4\,GHz of bandwidth.  Chapter \ref{chap:v4046sgr_structure} features an 
exquisite dataset from the SMA that combines observations taken by multiple configurations.  A major upgrade to 
the array is increasing the bandwidth to 8\,GHz, enabling more sensitive observations and larger spectral 
coverage.  To participate in VLBI experiments, the SMA can operate as a ``phased array'', essentially turning 
the interferometer into a single dish telescope.  Chapter \ref{chap:aphids} addresses a key step of how to 
integrate the SMA into millimeter VLBI observations such as those taken by the EHT.

Located 4450\,km away in Chile is the Atacama Large Millimeter Array (ALMA).  This facility began early
science operations in 2011 and has been commissioned in a staged fashion where each observing cycle offers more 
capabilities.  Once completed, ALMA will host an impressive 66 antennas, 64 of will be 12\,meter dishes and
the remaining smaller 7\,meter ones (see Figure \ref{fig:sma_alma}).  We used publickly available data from 
this instrument in Chapters \ref{chap:hd163296} and \ref{chap:radial_flows}.  A fully extended configuration 
will include baselines up to 16\,kilometer and the array will observe at wavelengths from 400\,microns to 
7\,millimeters.  The ALMA Phasing Project (APP) will enable this giant array to join VLBI experiments and 
fringes to APEX were found during a January 2015 test \citep{matthews15}.

@@

Both the SMA and ALMA are excellent observatories with which to study young stars.  Within a few hundred parsecs 
of the Sun there are multiple star forming regions and stellar moving groups.  Their relative proximity to us 
makes it possible to resolve physical structures in these regions and our great observatories have returned 
spectacular glimpses of the complicated processes that produce a star.  The canonical picture of star formation 
breaks this process into several stages \citep{shu87,mckee07}.  Picture first a large 
molecular cloud within which there are complicated sub structures: clumps and cores of 
concentrated gas and dust.  Over time these self-gravitating units will contract,  the material collapses into 
a disk to conserve angular momentum, and nascent stars ignite their thermonuclear engines.  Jets, winds, and 
outflows punctuate this process.  The system might also be perturbed by the swift pull of a nearby passing star 
or the constant presence of stellar partners forming a hierarchical system.  Whole fields, techniques, and 
technologies have been developed to study single chapters of this 
complicated story and ours begins near its end when only the disk and young star remain.

Over time, this disk of gas and dust will dissipate leaving trace debris: astroids and comets, dust and planets.
So while we may be at the end of the star's formation, this protoplanetary disk is the stage where planets are
born.  Their composition and structure provides the crucial initial conditions that must enable a diverse 
population of planets.  Theories of planet formation are still developing and require the consideration 
of physical processes acting on many physical scales and over a huge range of times.  As such, the
work to test and develop these ideas is quite varied, ranging from theoretical calculations to controlled 
experiments in the lab to the frontiers of space exploration.

@@

Over the past several decades, observations from interferometers operating at millimeter wavelengths have made
significant contributions to understanding the structure of these disks.  The early maps of the CO emission
made using 3 to 5 telescopes in an array at Owens Valley revealed bound, rotating disks with diameters 
similar to the dimensions of own solar system \citep{sargent87,weintraub89,koerner93}.  Emission from other 
molecules were then surveyed \citep{dutrey97} and studies of multiple isotopologues were used to constrain the 
two-dimensional structure of some disks \citep{dartois03,pietu07,qi11} aided by the development of 
chemical models \citep{aikawa99,aikawa01}.  Images of the dust also revealed interesting structures such as
giant holes centered around the star for a class of ``transition disks'' \citep{hughes07,brown09,andrews11} and 
evidence for binary interactions \citep[i.e.][]{akeson98}.

@@

There exists several techniques for constraining the structure of protoplanetary disks.  Some geometric 
properties such as size, inclination, and orientation can be calculated from an image.  The amount of dust 
in a disk can also be estimated by assuming a dust opacity and measuring the millimeter wave flux.  
Alternatively, one may try to solve the inverse problem: assume some disk model, calculate the resulting 
emission, and then compare the model's prediction to what is observed \citep[i.e.][]{dartois03,qi03}.  Often the 
density, temperature, and velocity profiles of the disks are assumed to 
follow power laws or some other parametric function.  There also exist some grids and models that try to 
calculate abundances in a self-consistent manner.  Additionally, researchers usually either try to constrain 
populations using 
surveys \citep[i.e.][]{andrews11,oberg11,harris12} or present detailed studies of single, 
interesting disks.  The work
presented in chapters \ref{chap:v4046sgr_structure} to \ref{chap:radial_flows} takes the latter approach and 
uses simple parametric 
models to investigate the structure of a choice few, but very interesting, protoplanetary disks.

V4046 Sgr is a system with two stars closely orbiting each other and around this tight binary is a bright
protoplanetary disk.  In chapter \ref{chap:v4046sgr_structure} we present SMA observations 
of this disk that reveal a large central cavity in the dust.  In addition, we show that the distribution dust 
does not follow that of the gas: the gas disk is large and extended while the 
millimeter sized dust grains are concentrated in a narrow ring.  Furthermore, thethe amount of dust left inside
the cavity is inconsistent with photoevaporative scenarios.  We argue 
that this configuration is the hallmark of an old disk that is being sculpted by at least one unobserved 
companion.

HD163296 also has a very large, gas rich disk, but no observed central cavity.  In chapter \ref{chap:hd163296} 
we use science verification data from ALMA to study the vertical structure of 
this disk.  We argue that resolved features in the images highlight how the temperature of the gas changes with
its distance from the disk mid-plane and explore how additional, nuanced aspects of the disk structure may affect
the data.  Chapter \ref{chap:radial_flows} then considers what impact radial motions of gas within transition 
disk holes may have upon spectral line emission.  We discuss several indicators of such radial flows along 
with possible sources of confusion such as warps in the disk.  One disk has provided intriguing
evidence of such flows \citep{casassus13,cassasus15} and future observations may provide others.

\begin{figure}[t]
\plotone{chapters/EHT-baselines.png}
\caption{A selection of EHT baselines viewed looking back from the galactic center where Sgr$^\ast$ lies.  The 
2015 campaign included sites 
from Hawaii, California, Arizona, Mexico, and Spain. ({\it image credit}: the EHT collaboration)}
\end{figure}

@@

The Event Horizon Telescope (EHT) observes supermassive black holes with baselines long enough to
to resolve event horizon scales of nearby targets.  These VLBI observations provide a 
unique glimpse into one of the most extreme environments in our universe.  Using telescopes in Hawaii, Arizona, 
and California, the collaboration has already detected structure at these scales for Sgr A$^\ast$ in the galactic 
center and the blaock hole at the center of the giant elliptical galaxy M87 
\citep{doeleman08,doeleman09}.  These data have provided indirect 
evidence for the existence of an event horizon \citep{broderick15} along with constraints on the degree of 
magnetization in the jet of M87 \citep{kino15} and the orientation of Sgr A$^\ast$ \citep{broderick11}.  EHT 
observations promise to provide tests of general relativity \citep{luminet79,johannsen10,bambi13,broderick14} as 
well as new studies of black hole time variability \citep{doeleman09} and accretion \citep{chan15}.

The bright future of the EHT is partly driven by its ability to add more telescopes and
incorporate new technologies.  The March 2015 observing campaign was a major milestone in
the EHT expansion with participation from the Large Millimeter Telescope (LMT) in Mexico, Combined Array for 
Research in Millimeter-wave Astornomy (CARMA) in California, the SMA and James Clerk Maxwell Telescope (JCMT) in
 Hawaii, SMT in Arizona and the 30m dish at Pico Valeta in Spain.  However, the data taken at the SMA must be 
processed before it
can be correlated with the rest of the array.  In chapter \ref{chap:aphids} we describe the APHIDS pipeline
which performs a sampling rate conversion using a cluster of Graphcs Processing Units (GPUs).  This software
is intended to be a long-term solution enabling VLBI with the SMA and has applications for other sites such 
as ALMA.

>>>>

Millimeter interferometry is intimately tied to the earlier development of techniques and technologies 
used at longer wavelengths.  The essential operation combines the signals from pairs of telescopes to produce a 
measure that is sensitive to some angular scale on the sky.  This scale shrinks as the distance
between the two elements grows.  By assembling an array with multiple 
telescopes, an observer can accumulate information about many spatial scales and reconstruct an image.  However,
the information contained in this image will be limited to the scales originally measured by the interferometer.
Images with the best fidelity come from arrays that have many elements spaced at a wide variety of 
distances and the highest resolution comes from the telescopes with the farthest separation.

@@

Each pair of antennas in an array forms a baseline vector (measured in wavelengths) with length $D_\lambda$ and 
components $(u,v,w)$.  The signal voltages, $V_1(t)$ and $V_2(t)$, measured by the individual 
elements are correlated against each other:
\begin{equation}
r(\tau) = \lim_{T \to \infty} \frac{1}{2T} \int_{-T}^T V_1(t) V_2^\ast(t-\tau) d\tau.
\end{equation}
This operation is often done on site with data streamed directly to a correlator and the resulting 
measurement, or visibility, is stored on disk.  However, for Very Long Baseline Interferometry (VLBI), the 
radio telescopes can be separated by entire continents and real-time correlation is not usually feasible.  For 
these experiments, the signal voltages must be recorded and the visibilities calculated at some later time.  The 
Event Horizon Telescope (EHT) is a leader in the developing field of millimeter VLBI and relies on existing 
observatories.  The difficulty of observing at these short wavelengths increases as the atmosphere becomes 
noiser, time variable, and more opaque.  Consequently, facilities that can operate at these wavelengths are a 
precious resource.

The Submillimeter Array \citep[SMA;][]{ho04} is one such world class array.  Commisioned in 2003, the SMA is 
located on the top of Maunakea on the Big Island of Hawaii at an altitude of 13,386 feet.
It is composed of eight 6 meter diameter dishes which, like many interferometers, can be moved into different 
configurations so that there are elements as close as 15 meters or as far as 500 meters (see Figure 
\ref{fig:sma_alma}).  Its recievers can operate at wavelengths from 430\,microns to 2\,millimeters and 
historically have provided up to 4\,GHz of bandwidth.  Chapter \ref{chap:v4046sgr_structure} features an 
exquisite dataset from the SMA that combines observations taken by multiple configurations.  A major upgrade to 
the array is increasing the bandwidth to 8\,GHz, enabling more sensitive observations and larger spectral 
coverage.  To participate in VLBI experiments, the SMA can operate as a ``phased array'', essentially turning 
the interferometer into a single dish telescope.  Chapter \ref{chap:aphids} addresses a key step of how to 
integrate the SMA into millimeter VLBI observations such as those taken by the EHT.

Located 4450\,km away in Chile is the Atacama Large Millimeter Array (ALMA).  This facility began early
science operations in 2011 and has been commissioned in a staged fashion where each observing cycle offers more 
capabilities.  Once completed, ALMA will host an impressive 66 antennas, 64 of will be 12\,meter dishes and
the remaining smaller 7\,meter ones (see Figure \ref{fig:sma_alma}).  We used publicly available data from 
this instrument in Chapters \ref{chap:hd163296} and \ref{chap:radial_flows}.  A fully extended configuration 
will include baselines up to 16\,kilometer and the array will observe at wavelengths from 400\,microns to 
7\,millimeters.  The ALMA Phasing Project (APP) will enable this giant array to join VLBI experiments and 
fringes to APEX were found during a January 2015 test \citep{matthews15}.

@@

Both the SMA and ALMA are excellent observatories with which to study young stars.  Within a few hundred parsecs 
of the Sun there are multiple star forming regions and stellar moving groups.  Their relative proximity to us 
makes it possible to resolve physical structures in these regions and our great observatories have returned 
spectacular glimpses of the complicated processes that produce a star.  The canonical picture of star formation 
breaks this process into several stages \citep{shu87,mckee07}.  Picture first a large 
molecular cloud made up of complex sub-structures: clumps and cores of 
concentrated gas and dust.  Over time these self-gravitating units will contract,  the material collapses into 
a disk to conserve angular momentum, and budding stars ignite their thermonuclear engines.  Jets, winds, and 
outflows punctuate this process.  The system might also be perturbed by the swift pull of a nearby passing star 
or the constant presence of stellar partners forming a hierarchical system.  Whole fields, techniques, and 
technologies have been developed to study single chapters of this 
complicated story and ours begins near its end when only the disk and young star remain.

Over time, this disk of gas and dust will dissipate leaving trace debris: astroids and comets, dust and planets.
So while we may be at the end of the star's formation, this protoplanetary disk is the stage where planets are
born.  Protoplanetary disk composition and structure provides the crucial initial conditions that must 
enable a diverse 
population of planets.  Theories of planet formation are still developing and require the consideration 
of physical processes acting on many physical scales and over a huge range of times.  As such, the
work to test and develop these ideas is quite varied, ranging from theoretical calculations to controlled 
experiments in the lab to the frontiers of space exploration.

@@

Over the past several decades, observations from interferometers operating at millimeter wavelengths have made
significant contributions to understanding the structure of these disks.  The early maps of the CO emission
made using 3 to 5 telescopes in an array at Owens Valley revealed bound, rotating disks with diameters 
similar to the dimensions of own solar system \citep{sargent87,weintraub89,koerner93}.  Emission from other 
molecules were then surveyed \citep{dutrey97} and studies of multiple isotopologues were used to constrain the 
two-dimensional structure of some disks \citep{dartois03,pietu07,qi11} aided by the development of 
chemical models \citep{aikawa99,aikawa01}.  Images of the dust also revealed interesting structures such as
giant holes centered around the star for a class of ``transition disks'' \citep{hughes07,brown09,andrews11} and 
evidence for binary interactions \citep{akeson98}.

@@

There exists several techniques for constraining the structure of protoplanetary disks.  Some geometric 
properties such as size, inclination, and orientation can be calculated from an image.  The amount of dust 
in a disk can also be estimated by assuming a dust opacity and measuring the millimeter wave flux.  
Alternatively, one may try to solve the inverse problem: assume some disk model, calculate the resulting 
emission, and then compare the model's prediction to what is observed \citep[i.e.][]{dartois03,qi03}.  Often the 
density, temperature, and velocity profiles of the disks are assumed to 
follow power laws or some other parametric function.  There also exist some model grids that provide more
physically motivated and self consistent disk structures \citep{dalessio05,woitke10}.  Additionally, researchers 
usually either try to constrain populations using 
surveys \citep[i.e.][]{andrews11,oberg11,harris12} or present detailed studies of single, 
interesting disks.  The work
presented in chapters \ref{chap:v4046sgr_structure} to \ref{chap:radial_flows} takes the latter approach and 
uses simple parametric 
models to investigate the structure of a choice few, but very interesting, protoplanetary disks.

V4046 Sgr is a system with two stars closely orbiting each other and around this tight binary is a bright
protoplanetary disk.  In Chapter \ref{chap:v4046sgr_structure} we present SMA observations 
of this disk that reveal a large central cavity in the dust.  In addition, we show that the dust 
distribution is significantly different than that of the gas: the gas disk is large and extended while the 
millimeter sized dust grains are concentrated in a narrow ring.  Furthermore, the amount of dust left inside
the cavity is inconsistent with photoevaporative scenarios.  We argue 
that this configuration is the hallmark of an old disk that is being sculpted by at least one unobserved 
companion.

HD163296 also has a very large, gas rich disk, but no central cavity has yet been observed.  In Chapter 
\ref{chap:hd163296} 
we use science verification data from ALMA to study the vertical structure of 
this disk.  We argue that certain resolved features in the images can be attributed to how the temperature of 
the gas changes with
its distance from the disk mid-plane and explore how additional, nuanced aspects of the disk structure may affect
the data.  Chapter \ref{chap:radial_flows} then considers what impact radial motions of gas within transition 
disk holes may have upon spectral line emission.  We discuss several indicators of such radial flows along 
with possible sources of confusion such as warps in the disk.  One disk has provided intriguing
evidence of such flows \citep{casassus13,casassus15} and future observations may be able to detect others.

\begin{figure}[t]
\plotone{chapters/EHT-baselines.png}
\caption{A selection of EHT baselines viewed looking back from the Galactic Center where Sgr$^\ast$ lies.  The 
2015 campaign included sites 
from Hawaii, California, Arizona, Mexico, and Spain. ({\it image credit}: the EHT collaboration)}
\label{fig:eht_baselines}
\end{figure}

@@

The Event Horizon Telescope (EHT) observes supermassive black holes with baselines long enough
to resolve event horizon scales of nearby targets.  These VLBI observations provide a 
unique glimpse into one of the most extreme environments in our universe.  Using telescopes in Hawaii, Arizona, 
and California, the collaboration has already detected structure at these scales for Sgr A$^\ast$ in the Galactic 
Center and the black hole at the center of the giant elliptical galaxy M87 
\citep{doeleman08,doeleman09}.  These data provide indirect 
evidence for the existence of an event horizon \citep{broderick15} along with constraints on the degree of 
magnetization in the jet of M87 \citep{kino15} and the orientation of Sgr A$^\ast$ \citep{broderick11}.  EHT 
observations promise to provide tests of general relativity \citep{luminet79,johannsen10,bambi13,broderick14} as 
well as new studies of black hole time variability \citep{doeleman09} and accretion \citep{chan15}.

The bright future of the EHT is partialy driven by its ability to add more telescopes and
incorporate new technologies.  The recent March 2015 observing campaign marked a major milestone in
the EHT expansion with participation from the Large Millimeter Telescope (LMT) in Mexico, Combined Array for 
Research in Millimeter-wave Astronomy (CARMA) in California, the SMA and James Clerk Maxwell Telescope (JCMT) in
Hawaii, SMT in Arizona and the 30m dish at Pico Valeta in Spain.  Figure \ref{fig:eht_baselines} shows the
impressive geographic span of the EHT including sites in Chile and at the South Pole that will soon be 
incorporated into the array.  However, the data taken at the SMA must be 
processed before it
can be correlated with the rest of the array.  In Chapter \ref{chap:aphids} we describe the APHIDS pipeline
which performs a sampling rate conversion using a cluster of Graphcs Processing Units (GPUs).  This software
is intended to be a long-term solution enabling VLBI with the SMA.

### 2015-09-03 frontmatter/abstract.tex
basic edits to intro and conclusion

Interferometric observations at millimeter wavelengths provide a precious, detailed view 
of certain astrophysical objects.  This thesis is composed of studies that both rely on and enable this 
technique to study the structure of planet-forming disks and eventually image the closest regions around 
super-massive black holes.  Young stars form out of a cloud of gas and dust that, before its eventual dissipation,
flattens to a disk.  However the disk population is diverse and recent high-resolution images have 
revealed a wide variety of interesting features.  To understand these observations we use detailed radiative 
transfer models to motivate various physical scenarios.  First we identify a set of traits in the disk around
V4046 Sgr that marks the coupled progression of the gas and dust distributions in the presence of at least one
embedded companion.  Next, we investigate how the vertical temperature structure of the disk can be spatially 
resolved and apply our framework to observations of the disk around HD163296.  Lastly, we show how the radial 
flow of gas may be imprinted in the data and question how this phenomenon might be distinguished from other 
scenarios such as warps or outflows.  The last chapter summarizes a solution on how to change the sampling 
rate of data taken at the SMA so that it may be used for VLBI campaigns.

>>>>

Interferometric observations at millimeter wavelengths provide a precious, detailed view 
of certain astrophysical objects.  This thesis is composed of studies that both rely on and enable this 
technique to study the structure of planet-forming disks and soon image the closest regions around 
super-massive black holes.  Young stars form out of a cloud of gas and dust that, before its eventual dissipation,
flattens to a disk.  However the disk population is diverse and recent high-resolution images have 
revealed a wide variety of interesting features.  To understand these observations we use detailed radiative 
transfer models to motivate various physical scenarios.  First we identify a set of traits in the disk around
V4046 Sgr that marks the coupled progression of the gas and dust distributions in the presence of at least one
embedded companion.  Next, we investigate how the vertical temperature structure of a disk can be spatially 
resolved and apply our framework to observations of the disk around HD163296.  Lastly, we show how large-scale
radial flows of gas may be observable and question how this phenomenon might be distinguished from other 
scenarios such as warps or outflows.  The last chapter summarizes the APHIDS project which changes the sampling 
rate of data taken at the SMA so that it may be used for VLBI campaigns.

### 2015-09-03 frontmatter/authorlist.tex
basic edits to intro and conclusion

\noindent The following authors contributed to Chapter 4: \newline
Lindy Blackburn, Rurik Primiani, Laura Vertatschitsch, and Andre Young.
}

>>>>

\noindent The following people contributed to Chapter 4: \newline
Lindy Blackburn, Geoff Crew, Sheperd Doeleman, Rurik Primiani, Mike Titus, Laura Vertatschitsch, 
Jonathan Weintroub, and Andre Young
}
